\section{Chapter 5: Nuclear and renewables}

\subsection{Nuclear}

The global nuclear plant fleet is 450 reactors producing around 11\% of global
electricity.

Long opposition was caused by linking nuclear power to nuclear winter
post-nuclear war stories.

\textit{
Nuclear energy is a result of splitting uranium 235, an isotope found in rock
at about 2–3\% of uranium, depending on the origin of the ore. Most of the
balance is non-fissile uranium, U-238. Each split uranium 235 atom releases
heat, atomic fragments, and one or two neutrons can cause other uranium 235
atoms to also split. A consistent rate of radioactive decay of uranium 235
occurs in the core of a nuclear plant. This heat is used to make steam, turn a
turbine, and spin a generator. The reactor core is kept cool usually by
circulating water in the tower. Passively cooled power plants, which are not
yet commercially in operation, propose using gravity, density, or some other
physical process to ensure that cores of reactors are kept cool without the
need for an external energy source. Most nuclear power plants are either
boiling water reactors or pressured water reactors. The major difference is in
how the heat is transferred in the process of making steam.
}

Top two producers of uranium ore are Canada and Australia.

\subsection{Wind}

Power = $0.5 \times \rho \times C \times v^3 \times A_{\text{Sweep}}$

\begin{itemize}
	\item $V^3$: meters per second -- doubling of the wind results in
	eightfold increase in power
	\item $\rho = 1.2kg \cdot m^{-3}$ (kilograms per cubic meter)
	\item $A_S$: $m^2$ -- area swept by the blade
	\item $C$ -- efficiency coefficient depending on the turbine, usually
	between 0.4 and 0.5
\end{itemize}
